---
title: "rNADEF Workbook"
author: North American Dendroecological Fieldweek <br> <i><small>Script by Stockton Maxwell, NADEF Co-Organizer; Github and Website by Chris Gentry; and data from Grant Harley, NADEF Co-Organizer,  Stockton Maxwell, Chris Gentry, and many others</small></i> <br>
output:
  html_notebook:
    df_print: paged
    rows.print: 10
    theme: cosmo
    highlight: breezedark
    number_sections: yes
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---
<style type="text/css">

h1.title {
  font-size: 40px;
  font-family: "Times New Roman", Times, serif;
  color: DarkGreen;
  text-align: Left;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
  font-size: 20px;
  font-family: "Times New Roman", Times, serif;
  color: DarkGreen;
  text-align: Left;
}
</style>

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

# Getting Started
## Libraries Used in this Document<br>
There are a number of useful libraries for manipulating and analyzing tree-ring data in **R**. In this document we will use the following tree-ring specific and general libraries.</br> 
<div class="col2">
<ul>
+ burnr
+ dendroTools
+ DendroSync
+ dplR
+ ggplot2
+ TRADER
+ treeclim
+ graphics
+ utils
</ul></div>

## Installing and Loading Libraries<br>
If you do not have one or many of these libraries installed you can add them by using the  ```install.libraries(x)``` function where *x* is the name of the package you desire to install. Start by loading all of the libraries:</br>

```{r libraries, message=FALSE, paged.print=FALSE}
library(burnr)
library(dendroTools)
library(DendroSync)
library(dplR)
library(ggplot2)
library(TRADER)
library(treeclim)
library(graphics)
library(utils)
```

All of these packages are cited in the [README](https://github.com/chrismgentry/rNADEF/blob/master/README.md) file available in the repository found at [rNADEF Workbook](https://github.com/chrismgentry/rNADEF)

# COFECHA in dplR
Using the script below, you will be able to read in raw ring width files for *COFECHA*-like analysis. To run your own data, change the ```fname```  to match the data in your ```data <- read.rwl(fname = ... , format = "auto")``` script. You can also examine summary statistics of your ring width file using the ```rwl.stats(x)``` function. Alternatively you can use the RWL file included in this project.

```{r Import Data and Analyze, echo=TRUE, message=FALSE, warning=FALSE}
grow.rwl <- read.tucson(fname = "Yellowstone_PSME_format.txt")
rwl.stats(grow.rwl)
```

You can plot this information to view the time series and/or skeleton plot of a series. To see the time series, use the function ```seg.plot(x)```. To view the skeleton plot of an individual series, use ```skel.plot(x[#])``` where # is the series/core you would like to view.

```{r Plot Time Series, echo=TRUE, fig.height=5, fig.width=10}
seg.plot(grow.rwl)
skel.plot(grow.rwl[1])
```

You can also examine the radii (mm) for each series, common time interval, and mean sensitivity of the rwl. While there isn't a direct function in *dplR* to calculate radii, you can use ```colSums()``` from base R to calculate the values. For the common interval and mean sensitivity you can use ```common.interval()``` and ```sens1()``` respectively.

```{r Radii-ComInt-MeanSen, echo=TRUE, fig.height=10, fig.width=10}
colSums(grow.rwl, na.rm = TRUE, dims = 1)
crn.common <- common.interval(grow.rwl, type=c("years"), make.plot=TRUE)
sens1(grow.rwl)
```

## Crossdating

Using functions from **dplR** you can also obtain a *COFECH*-like output with the ```corr.rwl.seg()``` function. You can also examine an individual series with ```series.rwl.plot()```. Please note that the options used in the examples below are specific to this project and might be different from your specific analysis.

```{r Crossdating, echo=TRUE, results='hide', fig.keep='all', fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
corr.rwl.seg(rwl = grow.rwl, seg.length = 50, bin.floor = 100, n = NULL, prewhiten = TRUE, pcrit = 0.05, biweight = TRUE, method = c("spearman"), make.plot = TRUE, label.cex = 1, floor.plus1 = FALSE, master = NULL)
```
To examine an individual series you will need to be able to identify the specific series in the function.
```{r Series Plot, echo=TRUE, results='hide', fig.keep='all', message=FALSE, warning=FALSE, fig.height=10, fig.width=10}
series.rwl.plot(grow.rwl, series = "BTK05A", series.yrs = as.numeric(names(series)), seg.length = 50, bin.floor = 100, n = NULL, prewhiten = TRUE, biweight = TRUE, floor.plus1 = FALSE)
```

Using ```interseries.cor``` will allow you to calculate the interseries correlation between each series and the master for the ring width file.

```{r Interseries Correlation, echo=TRUE}
interseries.cor(grow.rwl, n = NULL, prewhiten = TRUE, biweight = TRUE, method = "spearman")
```

To facilitate crossdating you can also use *dplR* to calculate marker rings with the ```pointer()``` function. 

```{r Marker Rings, echo=TRUE}
markers <- pointer(grow.rwl)
markers
```


## Spaghetti Plot of Raw Ring Widths

Frequently you will see a plot of the ring widths of all series on the same graph. These *spaghetti plots* can be created using the ```spag.plot()``` function in **dplR**.
```{r Spag Plot, echo=TRUE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
spag.plot(rwl = grow.rwl, zfac = 1, useRaster = FALSE, res = 300)
```

# Arstan in dplR
## Detrending

The standalone **ARSTAN** program allows you to develop site chronologies and interactively detrend or standardize a series. These options are also baked into **dplR**. To run interactive detrending you can use the ```i.detrend()``` function. Ihis allows you to explore curve fits for each tree ring series.
```{r iDetrending, include=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
grow.rwi.int <- i.detrend(rwl = grow.rwl, nyrs = NULL, f = 0.5,pos.slope = FALSE) #allows you to see a variety of fits
```
You will need to select a detrending method for each series.

To view the spaghetti plot for the detrended series you can once again use the ```spag.plot()``` function similar to above except substituting the raw ring width data for the detrended data.
```{r Detrend Spag Plot, echo=TRUE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
spag.plot(rwl = grow.rwi.int, zfac = 1, useRaster = FALSE, res = 300)
```

If you want to utilize a singular detrending method for all of your series, or you used interactive detrending to determine the best fit for your data, you can use the ```detrend()``` function instead of ```i.detrend()```. In this function you will choose one of the following detrending options with *method = *: "Spline", "ModNegExp", "Mean", "Ar", "Friedman", or "ModHugershoff".

```{r Detrending, echo=TRUE, message=FALSE, warning=FALSE}
grow.rwi <- detrend(rwl = grow.rwl, method = c("Spline"), nyrs = NULL, f = 0.5, pos.slope = FALSE)
grow.rwi
```

To examine the statistics for the entire chronology you can use ```rwi.stats()``` or ```rwi.stats.running()``` to use running statistics in order to adjust the time periods. See ```?rwi.stats.running``` for help on this function.

```{r Stats, echo=TRUE, message=FALSE, warning=FALSE}
rwi.stats(grow.rwi)
```

```{r RunStats, echo=TRUE, message=FALSE, warning=FALSE}
rwi.stats.running(grow.rwi)
```

## Generating Chronologies

The ARSTAN program generates a standardized chronology, an ARSTAN chronology, and a residual chronology. In *dplR* you can use the ```chron()``` function to build a mean value chronology from detrended ring widths produced from a ```detrend()``` function. 

Standardized chronology:
```{r Standardized, echo=TRUE, message=FALSE, warning=FALSE}
grow.crn <- chron(x = grow.rwi, prefix = "BTK", biweight = TRUE, prewhiten = FALSE)
grow.crn
```

Standard and Residual chronologies:
```{r Residual, echo=TRUE, message=FALSE, warning=FALSE}
grow.crn <- chron(x = grow.rwi, prefix = "BTK", biweight = TRUE, prewhiten = TRUE)
grow.crn
```

You can then plot the chronology using ```crn.plot()``` to view chronologies developed with ```chron()```.

```{r ChronPlot, echo=TRUE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
crn.plot(crn = grow.crn, add.spline = TRUE, nyrs = NULL, f = 0.5, crn.line.col='grey50', spline.line.col='red', samp.depth.col='grey90', samp.depth.border.col='grey80', crn.lwd=1, spline.lwd=2.0, abline.pos=1, abline.col='black', abline.lty=1,abline.lwd=1, xlab="Time", ylab="RWI")
```

For daily and monthly dendroclimatological analysis in the **dendroTools** library you can save the developed chronologies as *.csv* files using the ```write.csv(x)``` function:

```{r Export CRN, echo=TRUE}
btk_std <- grow.crn[1] #subset only year (already as.numeric) and index columns
```
To export the data use the following script: ```write.csv(btk_std, file = "BTK_std.csv")```

## Wavelet Transform
Wavelet Transform allows you to look at frequencies or temporal patterns in your *crn* for paleoclimatological analysis.

```{r Wavelet, echo=TRUE, fig.keep='all', fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
Years <- as.numeric(rownames(grow.crn))
rings <- grow.crn[, 1]
tubular <- morlet(y1 = rings, x1 = Years, p2 = 9, dj = 0.1,
                   siglvl = 0.99)
wavelet.plot(tubular, useRaster = NA)
```

# Treeclim

The **TreeClim** package allows for the assessment of growth-climate relationships similar to the older *DendroClim2002* software. To begin, you will need to load a chronology. You can use your own chronology or the chronology developed in the previous steps. With the script below, you can view a summary of the chronology you will be using for this analysis. Alternatively, you can change the name of the chronology to one you have generated.
```{r ChronSum, echo=TRUE, message=FALSE, warning=FALSE}
#summary of the chronology
summary(grow.crn)
```

Next you will need to load climate data. There are a number of methods for adding climate data but this example has one as a csv file in the project.

```{r Climate Data, echo=TRUE, message=FALSE, warning=FALSE}
climate <- read.csv("WY_yellowstone_climate.csv", header = TRUE)
names(climate)
summary(climate)
ym <- climate[,1:2] #pull year and month columns
var1 <- climate[3] #pull climate variables, 1 or 2 at a time to avoid N problems
clim <- data.frame(c(ym, var1)) #build climate data frame 
summary(clim)
```


## Response Function Analysis
*Modeled after Dendroclim2002* 
*Can produce "static", "moving", or "evolving" using the ```dynamic = ``` argument in the ```dcc``` script.*

```{r Response Function, echo=TRUE, message=FALSE, warning=FALSE}
resp <- dcc(chrono = grow.crn, climate = clim, selection = -5:10, method = "response", dynamic = "evolving", win_size = 35, win_offset = 1, start_last = TRUE, timespan = NULL, var_names = NULL, ci = 0.05, boot = "std", sb = FALSE) #this is the main function in treeclim
coef <- coef(resp) #model coefficients 
plot(resp) #plot the model coefficients
resp #show model results if you'd like
traceplot(resp, variables = c("PCP.prev.aug", "PCP.curr.jun", "PCP.curr.jul"), facet = FALSE) #shows correlations over time if moving or evolutionary selected
```

You can save the output using the following script: ```write.csv(coef, file = ("pcp_coef.csv"))```

Additionally you can write the plot to a file:
```
tiff("resp_coef.tiff", width = 8, height = 4, units = 'in', res = 300)
plot.new()
plot(resp)
title(main = "Climate", xlab = "Month")
dev.off()
```
## Split Calibration with 2 Climate Variables/Months
```{r 2 Variable Evaluation, echo=TRUE, fig.keep='all', fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
recon <- dcc(chrono = grow.crn, climate = clim, selection = 6:7, #use a selection with recon variable of interest - modifiers like .mean or .sum can be used to average across months
             method = "response", dynamic = "static", win_size = 35,
             win_offset = 1, start_last = TRUE, timespan = NULL, var_names =
             NULL, ci = 0.05, boot = "std", sb = FALSE)
recon_coef <- coef(recon)
plot(recon)
recon #show model results
#this does the evaluation
skillz <- skills(object = recon, target = .mean(1:2), model = "ols", calibration = "50%", timespan = NULL)
plot(skillz)
skillz #show model results
```

## Split Calibration with Single Variable
*Is your model time stable and does it verify?*
```{r 1 Variable Evaluation, echo=TRUE, fig.keep='all', fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
recon_dlm <- dlm(chrono = grow.crn, climate = clim, selection = 6, timespan = NULL, var_names = NULL,
    param_names = NULL, intercept = TRUE, scale = FALSE)
recon_coef <- coef(recon_dlm)
plot(recon_dlm)
recon_dlm #show model results
#this does the evaluation
skillz <- skills(object = recon_dlm, target = 6, model = "ols", calibration = "50%", timespan = NULL)
plot(skillz)
skillz #show model results
```

## Seasonal Correlation Analysis
*Orginal code developed by Dave Meko in Matlab*
```{r Seasonal Correlation, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
climate <- read.csv("WY_yellowstone_climate.csv", header = TRUE)
names(climate)
summary(climate)
ym <- climate[,1:2] #pull year and month columns
var1 <- climate[3:4] #pull climate variables, 1 or 2 at a time to avoid N problems
clim <- data.frame(c(ym, var1)) #build climate data frame 
summary(clim)
seas <- seascorr(grow.crn, climate, var_names = NULL, timespan = NULL, complete = 9,
         season_lengths = c(1, 3, 6), primary = 1, secondary = 2, ci = 0.05)#this is the main function
plot(seas)
seas
```

# TRADER Code for Growth Release Detection

**TRADER**, Tree Ring Analysis of Disturbance Events in R, is a package for disturbance reconstruction from tree-ring data. Analyses include Absolute Increase (Fraver & White 2005), Growth Averaging (Nowacki & Abrams 1997), Boundary Line (Black & Abrams 2003), and Splechtna (Splechtna, Gratzer & Black 2005) which combines growth averaging and boundary line techniques. 

To start you need to read in a rwl file to begin the analysis.

```{r Trader Data, echo=TRUE, message=FALSE, warning=FALSE}
thedata <- read.tucson('Yellowstone_PSME_format.txt')
```

For this example the *Growth Averaging*, ```growthAveragingALL()```, technique will be used with a the following paramenters: years averaged prior to release = 10 (*m1*), years averaged after release = 10, number of years between release events = 10, moderate relase = 0.25, major release = 0.50, with 5yrs of exceeding growth to be considered a release. Note, some of the analyses will create a large number of files to be created in your working directory/

```{r NowackiAbrams, echo=TRUE, message=FALSE, warning=FALSE}
growthAveragingALL(thedata, releases = NULL, m1 = 15, m2 = 15,buffer = 10, drawing = TRUE, criteria = 0.25, criteria2 = 0.50,prefix = "ga", gfun = mean, length = 5, storedev = jpeg)
```

```This script was not run as part of the workbook to avoid a large number of files created in the GitHub repository```

Similar to the script in a previous section, you can create a spaghetti plot of the data.

```{r TRADER SpagPlot, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
spag.plot(thedata, zfac = 1, useRaster = FALSE, res = 300)
thedata.raw.crn <- chron(thedata, prefix = "BTK", prewhiten=FALSE)
plot(thedata.raw.crn,abline.pos=NULL,ylab='mm',xlab='Year')
```

# Basal Area Increment calculation in *dplR*

Using the ```bai.out``` function you can use the raw ring width file to calculate basal area increment.
```{r BAI, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
grow.rwl <- read.rwl(fname = "wa091.rwl", format = "auto")
basal <- bai.out(grow.rwl, diam = NULL)
basal_p <- print(basal)
spag.plot(basal[1], zfac = 1, useRaster = FALSE, res = 300)
```

Similar to before, you can export the data using 
```write.csv(basal_p, "basal_bai.csv")```

# burnR

The **burnR** package creates composite fire history plots similar to the **FHX2** and **FHAES** programs. It uses *FHX* formatted files. An example of a FHX file is included in the repository.

After loading the ```burnr``` library, you can use ```read_fhx()``` to load data already formatted into the FHX format.

```{r Load FHX, echo=TRUE, message=FALSE, warning=FALSE}
Zion <- read_fhx('Zion.fhx')
Zion
```

Although each of the samples have an ID, we can add so site information in order to facet our plots. This will sort each of the samples into the appropriate category when facted with the ```plot_demograph``` function.

```{r Plot Facet FHX, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
Sites <- read.csv('ZionSiteIDs.csv')
Sites
facetplot <- plot_demograph(Zion, facet_group = Sites$SiteID, facet_id = Sites$series, plot_legend = TRUE)
print(facetplot)
```

Another option available in ```plot_demograph``` is a to create a plot of all samples with a composite plot beneath the plot of individuals. In addition to the plot and composite, you can also add annotations to highlight common fire dates.

```{r Plot Comp FHX, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
rugplot <- plot_demograph(Zion, composite_rug = TRUE, plot_legend = TRUE)
compositerug <- rugplot + 
  annotate('rect', xmin = 1721, xmax = 1723, ymin = 0, ymax = 21, alpha = 0.4) +   annotate('rect', xmin = 1734, xmax = 1736, ymin = 0, ymax = 21, alpha = 0.4) +   annotate('rect', xmin = 1748, xmax = 1750, ymin = 0, ymax = 21, alpha = 0.4) +   annotate('rect', xmin = 1777, xmax = 1779, ymin = 0, ymax = 21, alpha = 0.4) +   annotate('rect', xmin = 1793, xmax = 1795, ymin = 0, ymax = 21, alpha = 0.4) +   scale_x_continuous(limits=c(1450, 2005), breaks = seq(1450,2005,25)) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.2), 
  panel.grid.major = element_blank(), panel.grid.minor = element_blank())
print(compositerug)
```

# dendroTools
Alright, let's step up the game! Let's try some climate response with daily climate data because trees don't know what a month is.  You can get daily climate data for the US from the [PRISM Climate Group](http://www.prism.oregonstate.edu/).

Load and format the data prior to analysis.
```{r Formatting PCP Data, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
pcp_d <- read.csv("BTK_daily_pcp.csv", header = TRUE)
row.names(pcp_d) <- as.numeric(pcp_d$Year)
pcp_d <- pcp_d[,2:367] #subset only 1 to 366 days, not year column
crn_d <- read.csv("Btk_std.csv")
row.names(crn_d) <- as.numeric(crn_d$X) #set the years as row names
crn_d <- crn_d[2] #subset only growth index
head(crn_d) 
```
Take a look at the daily climate data
```{r Glimpse, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
glimpse_daily_data(env_data = pcp_d, tidy_env_data = FALSE, na.color = "white")
```

Analyze growth vs climate with fixed window width. This is set in the script using the ```fixed_width = ``` function.
```
fixed_width <- daily_response(response = crn_d, env_data = pcp_d,
               method = "cor", fixed_width = 60, row_names_subset = TRUE,
               remove_insignificant = TRUE, alpha = 0.05)
```               
```{r Fixed Width, echo=TRUE, fig.height=6, fig.keep='all', fig.width=10, message=FALSE, warning=FALSE}               
fixed_width$plot_extreme #creates a plot showing best correlated period
```

You can compare the response across two periods of analysis to assess time stability, or you can leave the subset to all the years.

```
btk_past <- daily_response(response = crn_d, env_data = pcp_d,
            method = "cor", lower_limit = 50, upper_limit = 70,
            row_names_subset = TRUE, previous_year = TRUE,
            remove_insignificant = TRUE, alpha = 0.05,
            plot_specific_window = 60, subset_years = c(1981, 1998))
```
```
btk_present <- daily_response(response = crn_d, env_data = pcp_d,
               method = "cor", lower_limit = 50, upper_limit = 70,
               row_names_subset = TRUE, previous_year = TRUE,
               remove_insignificant = TRUE, alpha = 0.05,
               plot_specific_window = 60, subset_years = c(1999, 2016))
```

Now you can plot the results.
```{r BTK Past Present Plots, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
btk_past$plot_heatmap
btk_present$plot_heatmap
btk_past$plot_specific #choose a specific window length to plot if you set this above
btk_present$plot_specific
```
## Climate Response on the Principle Components

Load data
```{r PCA DAta, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
data(example_proxies_individual)
data(LJ_daily_temperatures)
```

Example PCA - just create a data frame with multiple crns
```
example_PCA <- daily_response(response = example_proxies_individual, 
                              env_data = LJ_daily_temperatures, method = "lm", 
                              lower_limit = 60, upper_limit = 70,
                              row_names_subset = TRUE, remove_insignificant =
                              TRUE, alpha = 0.01, PCA_transformation = TRUE,
                              components_selection = "manual", N_components = 2)
```
Summary output and plot for the PCA analysis.
```{r PCA, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
summary(example_PCA$PCA_output)
```
```{r PCA Heat Map, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
example_PCA$plot_heatmap
```

## Quick common period reconstruction

Load the sample data.
```{r CP Recon Data, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
data(data_TRW)
data(KRE_daily_temperatures)
```
Reconstruction
```
example_reconstruction_lin <- daily_response(response = data_TRW, 
                              env_data = KRE_daily_temperatures, method = "lm",
                              metric = "r.squared", lower_limit = 30,
                              upper_limit = 40, row_names_subset = TRUE,
                              temporal_stability_check = "progressive",
                              cross_validation_type = "randomized", k = 3)
```
Reconstruction Plots
```{r Recon Plots, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
example_reconstruction_lin$plot_extreme
example_reconstruction_lin$temporal_stability
example_reconstruction_lin$cross_validation
example_reconstruction_lin$transfer_function
```
```{r Recon Models, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
linear_model <- lm(Optimized_return ~ TRW, data = example_reconstruction_lin$optimized_return)
reconstruction <- data.frame(predictions = predict(linear_model, newdata = data_TRW))
linear_model <- lm(Optimized_return ~ TRW, data = example_reconstruction_lin$optimized_return)
reconstruction <- data.frame(predictions = predict(linear_model, newdata = data_TRW))
plot(row.names(data_TRW), reconstruction$predictions, type = "l", xlab = "Year", ylab = "Mean temperature May 15 - Jun 27 [?C]")
```
## Monthly Analysis
load data
```{r Monthly Analysis Data, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
flow_d <- read.csv("BighornXavier_r.csv", header = TRUE)
row.names(flow_d) <- as.numeric(flow_d$year)
flow_d <- flow_d[,2:13]#subset only 1 to 366 days, not year column
```

Run analysis with split period
```
flow_past <- monthly_response(response = crn_d, env_data = flow_d,
             method = "cor", row_names_subset = TRUE, previous_year = TRUE,
             remove_insignificant = TRUE, alpha = 0.05, subset_years = c(1935,
             1976), aggregate_function = 'mean')
```
```
flow_present <- monthly_response(response = data_MVA, env_data =
                LJ_monthly_temperatures, method = "cor", row_names_subset =
                TRUE, alpha = 0.05, previous_year = TRUE, remove_insignificant =
                TRUE, subset_years = c(1977, 2016), aggregate_function = 'mean')
```
```{r Flow Plots, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
flow_past$plot_heatmap
flow_present$plot_heatmap
flow_past$plot_extreme
flow_present$plot_extreme
```
## Monthly analysis using a PCA of chronologies with crns

You still need to get crns in column format
```
flow_PCA <- monthly_response(response = example_proxies_individual,
            env_data = flow_d, method = "lm", row_names_subset = TRUE,
            remove_insignificant = TRUE, alpha = 0.01, PCA_transformation =
            TRUE, previous_year = TRUE, components_selection = "manual",
            N_components = 2)
```
```{r Monthly CRN Plots, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
summary(flow_PCA$PCA_output)
flow_PCA$plot_heatmap
flow_PCA$plot_extreme
```
# DendroSync

This package provides functions for the calculation and plotting of synchrony in tree growth from tree-ring width chronologies (TRW index).

Load Data
```{r DendroSync Data, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
data(conifersIP) #note the format of the data if you choose to use this analysis
head(conifersIP)
```

Calculate synchrony for null.model (broad evaluation, mBE) and homoscedastic variant of unstructured model (or full, mUN) for conifersIP data, and heteroscedastic variant for 1970-1999 period.

Fit the homoscedastic set of varcov models (mBE, mNE, mCS, mUN) using taxonomic grouping criteria (i.e. Species)
```{r ModHM, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
ModHm <- dendro.varcov(TRW ~ Code, varTime = "Year", varGroup = "Species", 
                       data = conifersIP, homoscedastic = TRUE)

summary(ModHm)# Class and length of list elements
```

Examine synchrony for mBE and mUN models
```{r Model Synchrony, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
sync(ModHm, modname = "mBE")
sync(ModHm, modname = "mUN")
```

Subset the data from 1970 to 1999.
```{r Model Subset, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
conif.30 <- conifersIP[conifersIP$Year>1969 & conifersIP$Year<2000,]
summary(conif.30$Year)
```

Fit the heteroscedastic set of variance covariance mixed models (mBE, mHeNE, mHeCS, mHeUN) using taxonomic grouping criteria (ie. Species)
```{r Mixed Models, echo=TRUE, fig.keep='all', fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
ModHt30 <- dendro.varcov(TRW ~ Code, varTime = "Year", varGroup = "Species", 
                         data = conif.30, homoscedastic = FALSE)
```
```{r Mixed Model Synchrony, echo=TRUE, message=FALSE, warning=FALSE}
sync(ModHt30, modname = "mBE")
sync(ModHt30, modname = "mHeUN")
```

-----------------------------------------------------
This workbook was created as an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code. The resulting *html* is used to display the static contents via GitHub.
